---
title: "R/Pharma ADaM Workshop"
author: 
  - Fanny Gautier Cytel
  - Ben Straub GSK
  - Edoardo Mancini Roche 
  - Sadchla Mascary Gilead
date: 2024-10-25
date-format: full
format: 
  revealjs:
    theme: solarized
editor: 
  markdown: 
    wrap: 72
---

```{r}
#| include: false
library(glue)
library(countdown)
library(link)

link::auto(keep_pkg_prefix = FALSE,
           keep_braces = FALSE)

# Package hexes
# Package hexes
metacore_img <- "https://raw.githubusercontent.com/atorus-research/metacore/master/man/figures/metacore.PNG"
metatools_img <- "https://raw.githubusercontent.com/pharmaverse/metatools/master/man/figures/metatools.png"
xportr_img <- "https://raw.githubusercontent.com/atorus-research/xportr/master/man/figures/logo.png"
admiral_img <- "https://raw.githubusercontent.com/pharmaverse/admiral/master/man/figures/logo.png"
pharmaverse_img <- "https://raw.githubusercontent.com/pharmaverse/pharmaverse-pkg/master/man/figures/banner.png"
dplyr_img <- "https://raw.githubusercontent.com/tidyverse/dplyr/master/man/figures/logo.png"
str_img <- "https://raw.githubusercontent.com/tidyverse/stringr/master/man/figures/logo.png"
haven_img <- "https://raw.githubusercontent.com/tidyverse/haven/master/man/figures/logo.png"
lub_img <- "https://raw.githubusercontent.com/tidyverse/lubridate/master/man/figures/logo.png"

# Set image sizes
img_bullet_size <- 80
img_right_size <- 150
img_center_size <- 300

```

## Objectives

::: incremental
-   By the end of this workshop you will have:
    -   Updated code for ADaMs `ADSL` and `ADVS` using `pharmaverse`
        packages, `tidyverse` packages, base R and Pinnacle21-like
        specs.
    -   Gained an understanding of how to use `pharmaverse` packages
    -   Gained an understanding of how to use resources for building
        ADaMs in `R`.
:::

## Play by Play

::: incremental
- ðŸ•™10:05 - 10:15 Introductions and Setup

- ðŸ•™10:15 - 10:30 Overview of pharmaverse `packages` and functions

- ðŸ•¥10:30 - 11:20 `ADSL`

- ðŸ•š11:20 - 11:30 Break

- ðŸ•¦11:30 - 12:30 `ADVS`

- ðŸ•12:30 - 1:00 Questions and Resources

:::

# Introductions and Setup

ðŸ•™10:05 - 10:15

## Who will help you achieve the objective!

> > -   `Fanny Gautier` - `presenting on ADVS` ðŸ“£
> > -   `Ben Straub` - `presenting on ADSL and pharmaverse packages` ðŸ“£
> > -   `Edoardo Mancini` - `Chat Support` ðŸ’»
> > -   `Sadchla Mascary` - `Chat Support` ðŸ’»

## Setting up your Workspace

::: incremental
-   **Preferred**:

    -   Sign up for a free [Posit Account](https://posit.cloud/)

    -   Share Link for R/Pharma: ADaM Workshop

        -   Everything loaded (data, packages, specs) and ready to go!

-   **Advanced:**
    -   Clone
        [Repo](https://github.com/pharmaverse/rpharma-2024-ADaM-workshop)
        and set up yourself
    -   install.packages(c(admiral, xportr, metatools, ...))
    
:::

## How we will "code" today

::: incremental
-   Fanny and I will walk you through coding `ADSL` and `ADVS`
    -   Lite Discussion on each function
    -   Fill in the blank!
    -   Occasional Zoom Poll
-   Important to move along quickly
    -   Please post questions to chat
    -   Full scripts are available
:::

# Overview of pharmaverse packages and functions

ðŸ•¥10:15 - 10:30

## The core pharmaverse packages

-   `metacore`
    `r glue("<img src={metacore_img} height='{img_bullet_size}'/>")`
-   `metatools`
    `r glue("<img src={metatools_img} height='{img_bullet_size}'/>")`
-   `xportr`
    `r glue("<img src={xportr_img} height='{img_bullet_size}'/>")`
-   `admiral`
    `r glue("<img src={admiral_img} height='{img_bullet_size}'/>")`

## The core tidyverse packages

-   `dplyr`
    `r glue("<img src={dplyr_img} height='{img_bullet_size}'/>")`
-   `stringr`
    `r glue("<img src={str_img} height='{img_bullet_size}'/>")`
-   `haven`
    `r glue("<img src={haven_img} height='{img_bullet_size}'/>")`
-   `lubridate`
    `r glue("<img src={lub_img} height='{img_bullet_size}'/>")`

## Hello pharmaverse Functions! {.smaller}

## Hello pharmaverse Functions! {.smaller}

| Package                     | Function                                                  | Purpose                                                       |
  |-------------------|-------------------------------|----------------------|
  | `r link::to_pkg("admiral")` | `r link::to_call("admiral::derive_vars_merged()")`        | `left_join` on steroids                                       |
  | `r link::to_pkg("admiral")` | `r link::to_call("admiral::derive_param_computed()")`     | Core function for wrappers used in ADVS                       |
  | `r link::to_pkg("admiral")` | `r link::to_call("admiral::convert_blanks_to_na()")`      | SAS blanks versus R `NAs`                                     |
  | `r link::to_pkg("admiral")` | `r link::to_call("admiral::derive_vars_dtm()")`           | Take a `---DTC` variable and turn it into a `---DTM` variable |
  | `r link::to_pkg("admiral")` | `r link::to_call("admiral::derive_vars_duration()")`      | Duration between timepoints                                   |
  | `r link::to_pkg("admiral")` | `r link::to_call("admiral::derive_vars_extreme_event()")` | ðŸ¤¯                                                            |
  
## Hello pharmaverse Functions! {.smaller}
  
  | Package                       | Function                                                   | Purpose                                 |
  |-------------------|-------------------------------|----------------------|
  | `r link::to_pkg("metatools")` | `r link::to_call("metatools::combine_supp()")`             | Join Parent and Supplementary Datasets  |
  | `r link::to_pkg("metatools")` | `r link::to_call("metatools::create_var_from_codelist()")` | Numeric Vars from Specs!                |
  | `r link::to_pkg("metatools")` | `r link::to_call("metatools::drop_unspec_vars()")`         | Easily drop variables not in Spec       |
  | `r link::to_pkg("xportr")`    | `r link::to_call("xportr::xportr_label()")`                | Apply labels from Spec                  |
  | `r link::to_pkg("xportr")`    | `r link::to_call("xportr::xportr_write()")`                | Write out an xpt file                   |

## Hello Specs!

-   Pinnacle21-like Specs

-   Not fit for purpose - just a helpful guide

-   What is in the the Specs:

    -   Two datasets, dataset labels, Keys

    -   Variables - labels, lengths, types, linked to method

    -   Codelists!!

    -   No VLMs

# `ADSL`

ðŸ•¥10:30 - 11:20

# `ADSL` Quizzes

## Check-in 1

Which `admiral` function can be used to turn a `---DTC` variable into a `---DTM` variable

::: incremental
a) `derive_vars_dtc()`
b) `derive_var_trtdurd()`
c) `derive_vars_dtm()`
d) `derive_vars_exsdtc()`
:::

Correct Answer: c) `derive_vars_dtm()`

## Check-in 2

How many arguments does `derive_vars_merged()` have?

Hint: Go to the documentation ?fcn_name or F1

a) 15 arguments
b) 4 argruments
c) 1 argrument
d) 10 arguments

## Check-in 3

Which function did I use to load the `adams_little_helpers.R`?

a) `load_all()`
b) `install.functions()`
c) `source()`
d) `callr()`

## Check-in 4

How many packages did I load using `library()` to create `ADSL` today?

a) 15
b) 4
c) 6
d) 10

# Break

ðŸ•¦11:20 - 11:30

# `ADVS` 

ðŸ•¦11:30 - 12:30

# `ADVS` Quizzes

## Check-in 1

Have a look to [{admiraldiscovery}](https://pharmaverse.github.io/admiraldiscovery/articles/reactable.html):
Which admiral function could be used to derive "BMI" parameter at any time point?

> > -   a) admiral::compute_bmi()
> > -   b) admiral::derive_param_bmi()
> > -   c) [derive_params_growth_age()](https://pharmaverse.github.io/admiralpeds/reference/derive_params_growth_age.html)

> -   Correct answer:
> -   **b) the wrapper function admiral::derive_param_bmi()**

> -   We could also use `r link::to_call("admiral::derive_param_computed()")`

## Check-in 2

Have a look to [{admiraldiscovery}](https://pharmaverse.github.io/admiraldiscovery/articles/reactable.html):
Which admiral wrapper function could be used to derive "BSA" parameter at any time point?

> > -   a) admiral::derive_param_bsa()
> > -   b) admiral::compute_bsa()

> -   Correct answer:
> -   **a) the wrapper function admiral::derive_param_bsa()**

> -   We could also use `r link::to_call("admiral::derive_param_computed()")`

## Check-in 3

With the help of [{admiraldiscovery}](https://pharmaverse.github.io/admiraldiscovery/articles/reactable.html):
Which admiral function could be used to derive ONTRTFL variable?

> > -   a) admiral::derive_var_trtemfl()
> > -   b) admiral::derive_var_trtdurd()
> > -   c) admiral::derive_var_ontrtfl()

> -   Correct answer:
> -   **c) the function admiral::derive_var_ontrtfl()**

## Check-in 4

Execute the derivation of PCHG on a subset of the input dataset: Post-baseline records only.
Which admiral functions will you use ? Get some hints on the [Reference](https://pharmaverse.github.io/admiral/reference/) page

> -   Correct answer:
> -   **admiral::restrict_derivation()**
> -   **admiral::derive_var_pchg()**

## Check-in 5

With the help of the [Reference](https://pharmaverse.github.io/admiral/reference/) page:
which function could be used to compute the Analysis Sequence Number (Numbering the Observations Within Each Group)?

> -   Correct answer:
> -   **admiral::derive_var_obs_number()**

## Questions and Resources

> -   [admiraldiscovery](https://pharmaverse.github.io/admiraldiscovery/)
> -   [pharmaverse examples](https://github.com/pharmaverse/examples)
> -   pgkdown sites for xportr, metatools, metacore and admiral
> -   Extension packages for admiral


## Closing Thoughts {.smaller}

> -   We were only able to show a small number of pharmaverse packages
>     today, but please delve deeper to explore the full breadth
> -   If you'd like to contribute to `pharmaverse`, check out options at
>     [pharmaverse.org](https://pharmaverse.org/contribute/)
> -   Finally, be sure to join our community on
>     [Slack!](https://join.slack.com/t/pharmaverse/shared_invite/zt-yv5atkr4-Np2ytJ6W_QKz_4Olo7Jo9A)

## Pacakges and Session Information

```{r}
#| class-output: hscroll

sessionInfo()
```

```{css, echo=FALSE}
.hscroll {
  overflow-x: auto;
  white-space: nowrap;
}
```
