---
title: "R/Pharma ADaM Workshop"
author: 
  - Fanny Gautier
  - Ben Straub
  - Edoardo Mancini  
format: revealjs
---

```{r}
#| include: false
library(glue)
library(countdown)

# Package hexes
metacore_img <- "https://raw.githubusercontent.com/atorus-research/metacore/master/man/figures/metacore.PNG"
metatools_img <- "https://raw.githubusercontent.com/pharmaverse/metatools/master/man/figures/metatools.png"
xportr_img <- "https://raw.githubusercontent.com/atorus-research/xportr/master/man/figures/logo.png"
admiral_img <- "https://raw.githubusercontent.com/pharmaverse/admiral/master/man/figures/logo.png"
pharmaverse_img <- "https://raw.githubusercontent.com/pharmaverse/pharmaverse-pkg/master/man/figures/banner.png"

# Set image sizes
img_bullet_size <- 80
img_right_size <- 150
img_center_size <- 300

```


## Objective

- By the end of this workshop you will have developed an `ADSL` and `ADVS`
dataset using pharmaverse packages and dataset specs???

## Play by Play

10:00 - 10:10 Introductions and Setup
10:10 - 10:20 Overview of different packages and functions - table?
10:20 - 11:20 ADSL
11:20 - 11:30 Break
11:30 - 12:30 ADVS
12:30 - 1:00 Questions/Wiggle Room/Resources

* Questions pushed into the Chat for Edoardo - (Edoardo can come off mute to address)
* Feedback on Workshop versus Feedback on Packages

## What's covered in the Workshop

- How to use metacore/metatools and ADaM Specs for `ADSL` and `ADVS` 
- How to use admiral and other packages to code `ADSL` and `ADVS`
- Using Specs to create numeric variables and order and finalize dataset
- Write out final datasets as xpts (lengths, types, labels) using specs using `xportr`
- ???
  

## Helpful Resources

- admiraldiscovery
- pgkdown sites for xportr, metatools, metacore and admiral
- ?? TODO: link these out
- pharmaverse slack
- pharmaverse examples
- Extension packages for admiral

## Some things to ponder...

* Blanks versus NAs
* Default Arguments and `NULL` Arguments
* Viewing mini-datasets
* How to find the functions (discovery versus package documentation)
* Outline in the Code

## Blanks in the code!

* What to expect?
* 

## Packages Walkthrough - brief intro!

>> - `metacore` `r glue("<img src={metacore_img} height='{img_bullet_size}'/>")`
>> - `metatools` `r glue("<img src={metatools_img} height='{img_bullet_size}'/>")`
>> - `xportr` `r glue("<img src={xportr_img} height='{img_bullet_size}'/>")`
>> - `admiral` `r glue("<img src={admiral_img} height='{img_bullet_size}'/>")`

## Packages Walkthrough - metacore

## Packages Walkthrough - metatools

## Packages Walkthrough - xportr

## Packages Walkthrough - admiral

## Packages Walkthrough - example usage


## ADSL Walkthrough

- We are going to code ADSL 
- Parts of the code are blank out and will require some quick research
- Don't worry if you get lost we will get you back on track


```{r}
countdown(minutes = 50, seconds = 00, 
          top = NULL, bottom = 0,
          left = NULL, right = 0,
          margin = "10%",
          font_size = "2.5em")
``` 

## ADVS Walkthrough

- We are going to code ADVS 
- Parts of the code are blank out and will require some quick research
- Don't worry if you get lost we will get you back on track

```{r}
countdown(minutes = 50, seconds = 00, 
          top = NULL, bottom = 0,
          left = NULL, right = 0,
          margin = "10%",
          font_size = "2.5em")
``` 


## `admiral` exercises `r glue("<img src={admiral_img} align='right' height='{img_right_size - 20}'/>")`


```{r}
#| eval: false
#| echo: true

library(admiral)
library(dplyr, warn.conflicts = FALSE)
library(pharmaversesdtm)
library(lubridate)
library(stringr)

data("dm")
data("ds")
data("ex")
data("ae")

dm <- convert_blanks_to_na(dm)
ds <- convert_blanks_to_na(ds)
ex <- convert_blanks_to_na(ex)
ae <- convert_blanks_to_na(ae)
```


## `admiral` exercises - 1 `r glue("<img src={admiral_img} align='right' height='{img_right_size-20}'/>")` {.smaller}

- Create a new dataset `ds_ext` with all of the `ds`, plus derive a numeric date version of `DS.DSSTDTC` without any imputation as `DSSTDT`.  
*Hint: `derive_vars_dt()`*


```{r}
countdown(minutes = 10, seconds = 00, 
          top = NULL, bottom = 0,
          left = NULL, right = 0,
          margin = "10%",
          font_size = "2.5em")
```  

## `admiral` exercises - 2 `r glue("<img src={admiral_img} align='right' height='{img_right_size-20}'/>")` {.smaller}

- Create a new dataset `adsl_1` with all of `dm`, plus derive an end of study date (`EOSDT`) using the `DSSTDT` from `ds_ext` where `DSCAT == "DISPOSITION EVENT" & DSDECOD != "SCREEN FAILURE"`.  
*Hint: `derive_vars_merged()`*


```{r}
countdown(minutes = 10, seconds = 00, 
          top = NULL, bottom = 0,
          left = NULL, right = 0,
          margin = "10%",
          font_size = "2.5em")
```  

::: aside
[Solution](https://pharmaverse.github.io/admiral/cran-release/articles/adsl.html#disposition ){.external target="_blank"} 
:::

## `admiral` exercises - 3 `r glue("<img src={admiral_img} align='right' height='{img_right_size-20}'/>")` {.smaller}

- Create a new dataset `adsl_2` with all of `adsl_1`, plus derive a safety population flag (`SAFFL`) as `Y` for any patient from `ex` with a record where `EXDOSE > 0`.  
*Hint: `derive_var_merged_exist_flag()`*

```{r}
countdown(minutes = 10, seconds = 00, 
          top = NULL, bottom = 0,
          left = NULL, right = 0,
          margin = "10%",
          font_size = "2.5em")
```  

::: aside
[Solution](https://pharmaverse.github.io/admiral/cran-release/articles/adsl.html#popflag){.external target="_blank"} 
:::

## `admiral` exercises - 4 `r glue("<img src={admiral_img} align='right' height='{img_right_size-20}'/>")` {.smaller}

- Create a new dataset `adsl_3` with all of `adsl_2`, plus derive a last known alive date (`LSTALVDT`) as the latest date of the following variables from `ae`: `AESTDTC` and `AEENDTC`, imputing missing day and month to the first.  
*Hint: `derive_var_extreme_dt()`*

```{r}
countdown(minutes = 10, seconds = 00, 
          top = NULL, bottom = 0,
          left = NULL, right = 0,
          margin = "10%",
          font_size = "2.5em")
```  

::: aside
[Solution](https://pharmaverse.github.io/admiral/cran-release/articles/adsl.html#lstalvdt){.external target="_blank"} 
:::

## `admiral` exercises - 5 `r glue("<img src={admiral_img} align='right' height='{img_right_size-20}'/>")` {.smaller}

- If time permits, explore [higher order functions](https://pharmaverse.github.io/admiral/cran-release/articles/higher_order.html) that give greater flexibility to all `admiral` functions

```{r}
countdown(minutes = 10, seconds = 00, 
          top = NULL, bottom = 0,
          left = NULL, right = 0,
          margin = "10%",
          font_size = "2.5em")
```  

## Closing Thoughts {.smaller}

> - Anyone can be become a `pharmaverse champion` as one of the criteria we look for in `pharmaverse` recommended packages is great user guides, as hopefully demonstrated wit the sites shown today
> - We were only able to show a small number of pharmaverse packages today, but please delve deeper to explore the full breadth
> - If you'd like to contribute to `pharmaverse`, check out options at [pharmaverse.org](https://pharmaverse.org/contribute/)
> - Finally, be sure to join our community on [Slack!](https://join.slack.com/t/pharmaverse/shared_invite/zt-yv5atkr4-Np2ytJ6W_QKz_4Olo7Jo9A)




